% !TEX root = ../paper.tex
\begin{abstract}
Database access logs are knowledge bases for many forms of database administration, from database performance tuning, to security auditing, to benchmark design, and many more. They record how the data was being used, combined, and/or manipulated in the history and support decision making(e.g. query recommendation, indexing etc) for the future. The knowledge can be captured from examining \textit{abstract syntax trees}(ASTs) of queries which encode \textit{correlation} among different data operations and ideally database administrators learns a distribution over all possible sub-trees of ASTs in the log. This distribution helps administrators seek sub-trees they might be interested(frequent, outlier, etc). Unfortunately, it is challenging for mining such distribution over large number of syntax tree structures. For scalability, these tree structures should be transformed into a structurally simpler \textit{representation}. However, for knowledge preserving, the original syntactic tree structures must be efficiently re-constructed by parsing the representation and it is non-trivial to balance between. In addition, when the database administrator is human, in order to prevent the human from being overwhelmed by the potentially huge number of parameters or \textit{verbosity} in the actual distribution, a succinct \textit{summary} distribution with reduced verbosity should be visualized instead. Parameter reduction inevitably makes summary deviates from the actual distribution and the definition of \textit{error} measuring such deviation should be properly designed and benchmarked. The trade-off curve of error versus verbosity should be measured and visualized from which the administrator can dynamically balance between. It is then crucial to `optimize' such curve by minimizing the summarization error under all possible verbosity levels. In this paper, we offer a systematic approach that addresses the series of mentioned challenging problems aimed at helping database administrators analyze database access logs fast, scalable and precise.
\end{abstract}